---
title             : "Analysis of Covid Cases in North Dakota and the US"
shorttitle        : "Covid Forecasts"

author: 
  - name          : "Kevin Thompson"

affiliation:
  - id            : ""
    institution   : "Southern Methodist University Data Science Program"

authornote: |
  This project was completed in fulfillment of an assignment for a time series
  course at Southern Methodist University, though I plan to expand the scope
  of this analysis beyond the project requirements in the near future.

abstract: |
  The SARS-coronavirus-2 pandemic ravages the United States both economically,
  socially, and physically. Rapid, tailored responses to regional spikes in 
  cases depend on our ability to forecast the future state of the pandemic.
  The inherently explosive and non-stationary behavior of a pandemic presents
  its own modelling challenges along with the many sources of measurement error
  that come from both individual and structural faults.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "covid-19, SARS-coronavirus-2, time series analysis"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : no
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "doc"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
```

```{r analysis-preferences}

```

# Related Work

Predicting the behavior of viruses from the outbreak stage to the pandemic stage
has been an active field of study since the SIR model was introduced
in @kermack. The standard SIR begins with the modeling assumptions that viruses 
behave as lifeforms do (i.e. birth, life, reproduction, death) and that all members 
of a fixed population fall into the trichotomy expressed in the name of the model.
At every point in time, each individual in the model is either susceptible,
infectious, or recovered (SIR). Moreover, there is a fixed transmission and recovery rate. The SIR also assumes that the compartments in the trichotomy are ordered and that members of the population can only move in one direction. If an individual goes from infectious to recovered, they cannot be infected again.
Along with the above assumptions, the standard SIR model is a system of three
non-linear ordinary differential equations:


\begin{align}
  \frac{dS}{dt} &= -\beta\frac{IS}{N} \\
  \frac{dI}{dt} &= \beta\frac{IS}{N}-\gamma{I} \\
  \frac{dR}{dt} &= \gamma{I}
\end{align}

where S denotes the susceptible subpopulation size, I the infected subpopulation
size, and R the recovered subpopulation size [@assim]. $\gamma$ and $\beta$ are the
unknown recover rate and transmission rate parameters, respectively.

# Methods

## Data

The models in this paper are fit using data from the Johns Hopkins University
Coronavirus Research Center (JHU). The data collection process used by JHU  

## Univariate Models

## Multivariate Models

## SIR Model 

### No Modelling of Residuals

### With Modelling of Residuals

Given that the conventional time series methods like ARUMA models assume that
time is discrete, we will have to discretize equations (1)-(3). Thankfully, this
just leads to first differences because the differential equations are only
first order:

\begin{align}
  \frac{dS}{dt} \approx S_{t+1} - S_{t} \\
  \frac{dI}{dt} \approx I_{t+1} - I_{t} \\
  \frac{dR}{dt} \approx R_{t+1} - I_{t}
\end{align}

Following @assim I will label S as the total population and R will denote the
number of confirmed cases.

## Participants

## Material

## Procedure

## Data analysis
We use @assim. 


# Results

# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
